---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nestedcats

<!-- badges: start -->
<!-- badges: end -->

The goal of nestedcats is to ...

Manipulate nested categorical data in a tidy way, building off of dplyr, tidygraph, and igraph

This package aims to make it possible to store and manipulate rich hierarchical category information and do split-apply-combine operations on these categories.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jameelalsalam/nestedcats")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(tidygraph))
suppressPackageStartupMessages(library(ggraph))

library(nestedcats)

cats <- tibble(name = c("1", "1a", "1a1", "1a2", "1b"))

cat_tree <- tibble::tibble(
  parent = c("1", "1", "1a", "1a"),
  child =  c("1a", "1b", "1a1", "1a2")
)

cat_g <- tbl_graph(cats, cat_tree)

ggraph_nestedcats_tree(cat_g)

```

cat_tree <- tibble::tibble(
  parent = c(NA_character, "1", "1", "1a", "1a"),
  child = c("1", "1a", "1b", "1a1", "1a2")
) %>% as_closure_tbl()

leaf_data <- tibble::tibble(
  cat = c("1b", "1a1", "1a2"),
  value = c(1, 2, 3)
)

leaf_data %>%
  group_by_cats(cat = c("1a", "1b"), along = cat_tree) %>%
  summarize(value = sum(value, na.rm = TRUE))

verify_covers(c("1a", "1b"), over = cat_tree)
verify_disjoint(c("1a", "1b"), along = cat_tree)
verify_disjoint_covers(c("1a", "1b"), along = cat_tree)

all.equal(coverage(c("1a", "1b"), along = cat_tree),
          coverage("1", along = cat_tree))



