---
title: "Representing Nested Category Relationships"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

To represent the potentially complicated nesting of categories, we will use a graph. In particular, the graph must be a tree or directed acyclic graph (DAG) such that the edges represent subset relationships.

The machinery for working with trees and graphs can be a bit unfamiliar (especially to me) so this article explores the different ways that the graph data structure can be constructed and represented inside of nested categories and/or partially ordered factor objects.


```{r setup}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(igraph))
library(nestedcats)

ghg_nodes_df <- read_csv(system.file("extdata", "ghg_cats_nodes.csv", package = "nestedcats"))
ghg_edges_df <- read_csv(system.file("extdata", "ghg_cats_edges.csv", package = "nestedcats"))

ghg_edges_df
```

The correct format for the `igraph::graph_from_data_frame` constructor is a two-column data frame, with symbolic node names in each, and then a second data frame with node attributes. That is, the edge_df can have integer or character values.


```{r}
ghg_graph <- graph_from_data_frame(ghg_edges_df, vertices = ghg_nodes_df)
ghg_graph
```

The key features are that:

 * the secondary vertices data frame have the same list of node names as are used in the first two columns of the edge df.
 * the order of the secondary vertices df becomes the underlying node numbers in the graph.

